---
title: "Practical Machine Learning Project"
author: "Kyle Maurice"
date: "December 23, 2015"
output: html_document
---

Setup the libraries that will be used
```{r}
library(caret)
library(dplyr)
library(doParallel)
```

Set the seed for the program and set up parallel processing
```{r}
set.seed(6540)
c1 <- makeCluster(2)
registerDoParallel(c1)
```

Load training and testing data sets making sure to handle missing data

```{r}
training <- read.csv("pml-training.csv", stringsAsFactors=FALSE, na.strings = c("NA", "#DIV/0!", ""))
testing_final <- read.csv("pml-testing.csv", stringsAsFactors=FALSE, na.strings = c("NA", "#DIV/0!", ""))
```

Remove columns that have sparse data or data that will not be useful in a model.  

```{r}
count.na <- function(x) {sum(is.na(x))}
temp <- training %>% summarize_each(funs(count.na))
temp <- temp[,temp[1,]<=1000]
colkeep<-names(temp)
training <- training %>% select(one_of(colkeep))
testing_final$classe <- "X"
testing_final <- testing_final %>% select(one_of(colkeep))
testing_final <- testing_final %>% select(-one_of(c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window")))
training <- training %>% select(-one_of(c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window")))
training$classe <- factor(training$classe)
```

Split training data into test and training sets, since there are no classifications for the data set labeled test.

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
training <- training[inTrain,]
testing <- training[-inTrain,]

```

Set model parameters and train the random forest model. The mtry variable was set to 3 in order to speed up training.

```{r, cache=TRUE}
proj_model <- train(classe~., data = training, method="rf", tuneGrid=data.frame(mtry=3))
```

Summarize Model
```{r}
proj_model
```

Evaluate prediction on the reserved testing set. 
```{r}
testPredict <- predict(proj_model$finalModel, newdata=testing)
confusionMatrix(testPredict, testing$classe)
```

Cleanup
```{r}
stopCluster(c1)
```
